---
output: pdf_document
params: 
    set_title: "My Title!"
---


---
title: `r id`
---


| Type       | Your Average                     | Total Average                          |
|------------|----------------------------------|----------------------------------------|
| CALORIES   | `r round(mean(subgroup$KCAL),2)` | `r round(mean(map$KCAL,na.rm=TRUE),2)` |
| PROTEIN    | `r round(mean(subgroup$PROT),2)` | `r round(mean(map$PROT,na.rm=TRUE),2)` |
| TOTAL FAT  | `r round(mean(subgroup$TFAT),2)` | `r round(mean(map$TFAT,na.rm=TRUE),2)` |
| CARBS      | `r round(mean(subgroup$CARB),2)` | `r round(mean(map$CARB,na.rm=TRUE),2)` |
| FIBER      | `r round(mean(subgroup$FIBE),2)` | `r round(mean(map$FIBE,na.rm=TRUE),2)` |



##MicroNutrients

```{r, echo=FALSE}
dfm <- melt(subgroup[,c('StudyDayNo','PROT','TFAT','FIBE','CAFF','FIBE','SUGR')],id.vars = 1)
ggplot(dfm,aes(x = StudyDayNo,y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "stack") + scale_x_discrete(labels = abbreviate)
```

##Daily Calorie Intake
```{r, echo=FALSE}

p <- ggplot(subgroup,aes(x = StudyDayNo,y = KCAL)) + 
     geom_bar(aes(fill = StudyDayNo),stat = "identity",position = "dodge") + scale_x_discrete(labels = abbreviate)
p+guides(fill=FALSE)
```