---
output: 
  pdf_document: 
    latex_engine: xelatex
mainfont: Calibri
fig_caption: true
params: 
    set_title: "My Title!"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---
---
title: "Microbiome Diet Study Participant Report `r id`"
---



Thank you for participating in the Knights Lab's citizen science project! 

We have made you a personalized report from your samples and dietary data. Included in this report is information about your daily dietary intake and your daily microbiome variation. This report is not intended to provide medical advice or to be used to diagnose or treat disease. If you have any questions about your report, or you would like access to your raw sequence data, please contact the study coordinator Abby at cole0463@umn.edu.

#Your Dietary Intake Data

This first table and Figure 1 below show your average macronutrient intake as a percentage of calories and the typical macronutrient intake ranges for US adults:

| Type                | Average Macronutrient Intake                               | Typical Intake Range        |
|---------------------|------------------------------------------------------------:|-----------------------------:|
| CALORIES (kcal)     | `r round(mean(subnutr$KCAL),2)`                            |   1800 - 3200               |
| PROTEIN (%)         | `r round((mean(subnutr$PROT)/macro_totals)*100,2)`%        |   10 - 35%                  | 
| TOTAL FAT (%)       | `r round((mean(subnutr$TFAT)/macro_totals)*100,2)`%        |   20 - 35%                  |  
| CARBS (%)           | `r round((mean(subnutr$CARB)/macro_totals)*100,2)`%        |   45 - 65%                  |   

                                

Based on your height, weight, gender, age, and self-reported activity level you need approximately  `r signif(submap$TEE[1], 2)` kcal per day to maintain your current weight. During the study your weight change was `r submap$Weight.Change[1]` kg. If you didn't loose weight during the study, but your average intake was well below your estimated energy needs, it's possible that you slightly under-reported your dietary intake. 

##Figure 1:
```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.height=2.5, fig.width=6, fig.align="center"}


long_plot <- subnutr %>% select(X.SampleID, KCAL, PROT, TFAT, CARB, UserName)
long_plot$PROT <- long_plot$PROT*4/long_plot$KCAL
long_plot$TFAT <- long_plot$TFAT*9/long_plot$KCAL
long_plot$CARB <- long_plot$CARB*4/long_plot$KCAL


long_plot <- melt(long_plot, id.vars = c("X.SampleID", "UserName", "KCAL"), 
                  measure.vars = c("PROT", "TFAT", "CARB"), 
                  variable.name = "Macro", value.name = "PCT")


#instantiate color brewer and axis tick mark amount
colpal <- colorRampPalette(brewer.pal(9,"Set3"))(3)
colpal <- sample(colpal,3,replace = FALSE)
tick_label <- seq(1:17)

# plot kcal filled with % protein, fat, and carb
ggplot(long_plot, aes(x = X.SampleID, y = KCAL*PCT, fill = Macro)) +
  geom_bar(stat = "identity") +
  theme_classic() +
  scale_fill_manual(values=colpal, labels = c("Protein", "Total Fat", "Carbohydrate"))+
  #scale_fill_brewer(palette = colpal) +
  scale_x_discrete(labels=c(tick_label))+
  labs(fill = "Macronutrient\nDistribution", y = "Daily reported intake (Kcal)", x = "Study Day Number") +
  theme(
        #axis.title.x =element_text(family = "Helvetica"), axis.title.y =element_text(family = "Helvetica"),
        plot.title = element_text(hjust = 0.5))

``` 

**Figure 1** shows the day to day variation in your consumption of macronutrients viewed as a percentage of total kilocalorie intake.

Please note, it is common for people who follow some diets, like the low carbohydrate diet, to have intake ranges that differ from the typical intake ranges. The next table will show if your diet is nutritionally adequate.


\newpage
This second table shows your average daily intake of key micronutrients from food. Also shown are the average intake levels for all of our study participants and the recommended dietary allowance levels by gender. If you take a vitamin and mineral supplement, your intake from that supplement is not reflected here:


| Nutrient            | Your Average           | Study Average           |Recommended Dietary Allowance (Male/Female)|
|:--------------------|-----------------------:|------------------------:|:-----------------------------------------:|          
| FOLATE (mcg)        | `r round(mean(subnutr$FOLA),2)` | `r round(mean(nutrition_table$FOLA,na.rm=TRUE),2)` |                 400                      |
| CHOLINE (mg)        | `r round(mean(subnutr$CHOLN),2)`| `r round(mean(nutrition_table$CHOLN,na.rm=TRUE),2)`|                550 / 425                 |
| CALCIUM (mg)        | `r round(mean(subnutr$CALC),2)` | `r round(mean(nutrition_table$CALC,na.rm=TRUE),2)` |                 1000 *                   |
| SODIUM (mg)         | `r round(mean(subnutr$SODI),2)` | `r round(mean(nutrition_table$SODI,na.rm=TRUE),2)` |                 1500 *                   |
| POTASSIUM (mg)      | `r round(mean(subnutr$POTA),2)` | `r round(mean(nutrition_table$POTA,na.rm=TRUE),2)` |                 4700                     | 
| MAGNESIUM (mg)      | `r round(mean(subnutr$MAGN),2)` | `r round(mean(nutrition_table$MAGN,na.rm=TRUE),2)` |               400 / 310 *                | 
| IRON (mg)           | `r round(mean(subnutr$IRON),2)` | `r round(mean(nutrition_table$IRON,na.rm=TRUE),2)` |                8 / 18  *                 |
| SELENIUM (mcg)      | `r round(mean(subnutr$SELE),2)` | `r round(mean(nutrition_table$SELE,na.rm=TRUE),2)` |                 55                       |
| ZINC (mcg)          | `r round(mean(subnutr$ZINC),2)` | `r round(mean(nutrition_table$ZINC,na.rm=TRUE),2)` |                11 / 8  *                 |
| VITAMIN B12 (mcg)   | `r round(mean(subnutr$VB12),2)` | `r round(mean(nutrition_table$VB12,na.rm=TRUE),2)` |                 2.4                      |
| VITAMIN A (mcg)     | `r round(mean(subnutr$VARA),2)` | `r round(mean(nutrition_table$VARA,na.rm=TRUE),2)` |               900 / 700                  |
| VITAMIN C (mg)      | `r round(mean(subnutr$VC),2)`   | `r round(mean(nutrition_table$VC,na.rm=TRUE),2)`   |                90 / 75                   |
| VITAMIN D (mcg)     | `r round(mean(subnutr$VITD),2)` | `r round(mean(nutrition_table$VITD,na.rm=TRUE),2)` |                 15 *                     |
| VITAMIN E (mg)      | `r round(mean(subnutr$ATOC),2)` | `r round(mean(nutrition_table$ATOC,na.rm=TRUE),2)` |                 15                       |
| VITAMIN K (ug)      | `r round(mean(subnutr$VK),2)`   | `r round(mean(nutrition_table$VK,na.rm=TRUE),2)`   |               120 / 90                   |


*Values marked with an asterisk are the recommended dietary allowance or adequate intake values for adults between 18 and 30 years old. These values are the daily intake level that is sufficient to meet the needs of 97-98% of the population. If you are older than 30, your recommended dietary allowance may be higher or lower than what's shown here. You can visit https://ods.od.nih.gov/Health_Information/Dietary_Reference_Intakes.aspx for a complete list of dietary reference intakes by age and gender.

If your intake of a vitamin or mineral falls below the recommended level for your gender you should consider increasing your intake of foods that are a good source of that vitamin or mineral. For example, the study average intake of vitamin D falls below the recommended level of 15 mcg. So, many of the participants in our study could benefit by eating more vitamin D containing foods like fatty fish (such as salmon, tuna, and mackerel), fortified milk products, beef liver, egg yolks, and mushrooms. 


\newpage
##Figure 2:
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=7}

#instantiate color brewer and axis tick mark amounts
colpal<- colorRampPalette(brewer.pal(9,"Set3"))(37)
colpal<- sample(colpal,37, replace = FALSE)
tick_label <- seq(1:17) 

#Study average for individual's fiber intake
fiber_average <-round(mean(subnutr$FIBE),2)
fiber_rec <- round(14*mean(subnutr$KCAL)/1000)

p <- ggplot(subnutr,aes(x = StudyDayNo,y = FIBE)) + 
     geom_bar(aes(fill = colpal[1]),stat = "identity",position = "dodge") + scale_x_discrete(labels = abbreviate)+
      geom_hline(size=.75,col="red",yintercept = fiber_rec) + #Recommended fiber level
      geom_hline(size=.75,col="gold",yintercept = fiber_average) +  #Recommended fiber level
    annotate(geom="text", label="Recommended Intake Level",x=15,y=fiber_rec, vjust=-1, size = 3) + 
   annotate(geom="text", label="Your Average Intake Level",x=15,y=fiber_average,  vjust=-1, size =3) 

p + guides(fill=F) +
  theme_classic(base_family = "Helvetica")+
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_discrete(labels=c(tick_label))+
  scale_fill_manual(values = colpal)+
  labs(x="Study Day Number", y= "Fiber in grams")+
  ggtitle("Daily Fiber Intake")

```

**Figure 2** shows your fiber intake each day throughout the study. You had an average fiber consumption of `r format(signif(mean(subnutr$FIBE),2),scientific=FALSE)` grams per day. The red line in this plot shows the recommended level of fiber consumption in grams, while the gold line shows your average daily fiber intake. The current recommendation for fiber intake is to consume 14g per 1000kcal. We calculated your fiber needs to be `r fiber_rec` grams per day.

Fiber is an important energy source for the bacteria that live within your microbiome. Meeting the recommended intake of fiber may help to support a healthy microbiome. If your average fiber intake falls below the recommended intake level, you should consider increasing your intake of fiber by eating more fibrous vegetables and/or increasing your intake of whole grains.


##Figure 3:
```{r, warning=FALSE, message=FALSE, echo=FALSE}

# make ggplot bar chart of top 10 most abundant species per day
#melt sub_foodsp to get our dataframe in the long format for future usage (differs from melted_food)
  melt_food<- melt(sub_foodsp)
  
  #merge to get access to Day var (differs from merged_food)
  merge_food<- merge(x=melt_food, y=map, by.x = "variable", by.y= "X.SampleID", all.x=TRUE)
  
  #rename columns to get rid of the .x following UserName and StudyDayNo
  
  colnames(merge_food)[4] <- "UserName" #Change col name from UserName.x to UserName - compatibility purposes
  colnames(merge_food)[5] <- "StudyDayNo" #Change col name from StudyDayNo.x to StudyDayNo - compatibility purposes
  
  #convert our dataframe species (rn) column to a character
  merge_food$rn <- as.character(merge_food$rn) 
  
  #series of gsub commands meant to neaten and clarify legend content
  merge_food$rn <- gsub(".*s__", "", merge_food$rn)
  merge_food$rn <- gsub("\\[", "",merge_food$rn)
  merge_food$rn <- gsub("\\]", "",merge_food$rn)
  merge_food$rn <- gsub("_", " ",merge_food$rn)
  merge_food$rn <- gsub("L1", "",merge_food$rn)
  
  
  #instantiate color brewer and axes tick mark amount
  colpal <- colorRampPalette(brewer.pal(8,"Set3"))
  colpal <- sample(colpal(length(unique(merge_food$rn))))
  
  #made StudyDayNo a factor in the map, so don't need to use the tick labels

  
ggplot(merge_food, aes(x = StudyDayNo, y = value, fill = rn)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(drop = FALSE) +
   scale_fill_manual(values =colpal)+
  theme_classic(base_family = "Helvetica") +
  theme(axis.text.x = element_text( hjust = 0.5, family = "Helvetica"),
        plot.title =  element_text(hjust = 0.5, family = "Helvetica"),
        strip.background = element_rect(color = "grey")) +
  guides(fill = guide_legend(reverse = FALSE, 
                             keywidth = 1, 
                             keyheight = 1, 
                             ncol = 1,
                             title= "Food Group")) +
 labs(y="Relative Abundance",x="Study Day Number")+
  ggtitle("Food groups consumed each day")

```

**Figure 3** shows the relative abundance of the major food groups you consumed on a day to day basis during the study, excluding water. Each color is representative of a unique food group as annotated on the figure legend. Please note that these food groups are displayed by their weight in grams, so the plot tends to over-represent the contribution of liquid foods and foods with high water content to your diet. You may notice that the largest food component each day is called "Sugars, sweets and beverages". For most participants in this study, the size of that bar is driven by intake of coffee and tea, not sugar and candy.


\newpage
#Your Microbiome Composition Data

The next series of plots show information about the composition of your microbiome.  

We sequenced the DNA of the microbes that live within your gut from the daily stool samples you provided. We used a type of sequencing called Shallow-shotgun sequencing coupled with programs developed in the Knights's lab to identify the species of bacteria that live inside you.


##Figure 4:
```{r, warning=FALSE, message=FALSE, echo=FALSE}
# make ggplot bar chart of top 10 most abundant species per day
  
#melt subtaxasp to get our dataframe in the long format for future usage
  
  meltdf<- melt(subtaxasp)
  
  #merge to get access to Day var
  mergedf<- merge(x=meltdf, y=map, by.x = "variable", by.y= "X.SampleID", all.x=TRUE)
  
  #convert our dataframe species (rn) column to a character
  mergedf$rn <- as.character(mergedf$rn) 
  
  #series of gsub commands meant to neaten and clarify legend content
  mergedf$rn <- gsub(".*s__", "", mergedf$rn)
  mergedf$rn <- gsub("\\[", "",mergedf$rn)
  mergedf$rn <- gsub("\\]", "",mergedf$rn)
  mergedf$rn <- gsub("_", " ",mergedf$rn)
  
  
  #create small abundance category
  mergedf$rn[mergedf$value < 0.025] <- "<2.5% abundance"
  mergedf <- mergedf %>% group_by(StudyDayNo, rn) %>% summarise(newvalue = sum(value))
  
  
  #instantiate color brewer and axes tick mark amount
  colpal<- colorRampPalette(brewer.pal(12,"Set3"))
  colpal<- sample(colpal(length(unique(mergedf$rn))))

  
ggplot(mergedf, aes(x = StudyDayNo, y = newvalue, fill = rn)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(drop = FALSE) + 
  scale_fill_manual(values =colpal)+
  theme_classic(base_family = "Helvetica") +
  theme(axis.text.x = element_text( hjust = 0.5, family = "Helvetica"),
        plot.title =  element_text(hjust = 0.5, family = "Helvetica"),
        strip.background = element_rect(color = "grey")) +
  guides(fill = guide_legend(reverse = FALSE, 
                             keywidth = 1, 
                             keyheight = 1, 
                             ncol = 1,
                             title= "Bacterial Species")) +
 labs(y="Relative Abundance",x="Study Day Number")+
  ggtitle("Bacterial species in your gut each day")


```

**Figure 4** shows the most abundant bacterial species within your gut per each day of the study. The "<2.5% abundance" column represents a sum of bacterial species that individually account for less than 2.5% of the bacterial population within your gut microbiome. If you are missing microbiome data for any day on this plot it's either because you did not submit a sample for that day, your sample wasn't able to be sequenced, or we did not get enough sequence data from your sample to reliably use it for data analysis.



##Figure 5:
```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.width=6.5} 
 #transpose
   subtaxaalpha <- t(subtaxaalpha)
   taxalpha <- t(taxalpha)
 
  #calculate alpha diversity and assign to variable alphad
   alphad <- diversity(subtaxaalpha, index = "shannon", MARGIN = 1, base = exp(1))
   alphad <- as.data.frame(alphad)
   alphad <- rownames_to_column(alphad, var = "X.SampleID")
   
   #calculate study alpha diversity and assign to variable studyalpha
   studyalpha <- diversity(taxalpha, index = "shannon", MARGIN = 1, base = exp(1))
   studyalpha <- as.data.frame(studyalpha)
   studyalpha <- rownames_to_column(studyalpha, var = "X.SampleID")
   
   #merge melted df with map to get access to StudyDay variable
   merged_alpha<- merge(x=alphad, y=map, all.x=TRUE)
   
   # make studydayno a continuous variable for plotting
   merged_alpha$StudyDayNo <- as.numeric(merged_alpha$StudyDayNo)

   # Find statistical artifacts for (shannon) alpha diversity for all subjects
   med <- quantile(studyalpha$studyalpha)[3]
   q1 <- quantile(studyalpha$studyalpha)[2] #25th percentile
   q3 <- quantile(studyalpha$studyalpha)[4] #75th percentile

   #make gg plot line graph of subject alpha diversity per day
ggplot(merged_alpha, aes(x=StudyDayNo, y=alphad)) +
  scale_x_continuous(breaks = c(tick_label)) +
  theme_classic() +
  geom_smooth(color = "Black", fill = "#8DD3C7") +
  geom_point() +
  geom_hline(size=.75, color="grey", yintercept = med)+
  geom_hline(size=.75, color="#E3D5DC", yintercept = q1, linetype = "dashed")+
  geom_hline(size=.75, color="#E3D5DC", yintercept = q3, linetype = "dashed")+
  annotate(geom="text", label="Median",x=15,y=med, vjust=-1, size = 3) + 
  annotate(geom="text", label="25th percentile",x=15,y=q1, vjust=-1, size = 3) + 
  annotate(geom="text", label="75th percentile",x=15,y=q3, vjust=-1, size = 3) + 
  labs(x= "Study Day Number",y="Alpha Diversity (Shannon)")+
  guides(fill=FALSE)+
  ggtitle("Alpha diversity of your gut microbiome")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
  
```

**Figure 5** shows how the bacterial diversity exhibited within your gut changes on a daily basis. The measure of diversity shown here is called the Shannon index of alpha diversity. The Shannon index accounts for both the abundance and evenness of bacterial species present within the gut microbiome. A higher Shannon index typically indicates a more diverse microbiome community.

\newpage
##Figure 6:

```{r,fig.width=10, fig.height=10, echo=FALSE, warning = FALSE, message=FALSE}
#create PCOA plot of beta diversities of pertinent subjects

betad<-vegdist(betataxa, method="bray", binary=FALSE, diag=FALSE, upper=FALSE,
        na.rm = FALSE) 

#betad as matrix 
betad <- as.matrix(dist(betad))

#column name is "IDs"
IDs <-colnames(betad)

# Run the pcoa() function on the beta diversity table, and store the vectors generated as a dataframe 
PCOA <- data.frame(pcoa(betad)$vectors)

#create vector with placeholders
new_names <- rep("",ncol(PCOA))

#create for loop which gives column names (currently named "axis") the name of PC1, PC2, PC3.. so forth
#Each column for the PCOA matrix is a new PC (number).

for(i in 1:ncol(PCOA)){
  new_names[i] <- paste("PC",i,sep="")  
}

#code above uses a for loop - from the range of 1 to the number of columns in the PCOA
#matrix, the for loop accesses the new name variable and names it PC + i (which is 1,2,3  depending on column
#as indicated by for loop - for loop increases in increments of 1 (i goes up by 1) and goes thru columns)

#new_names (PC1, PC2, etc..) replace the current colnames
names(PCOA) <- new_names

#creates column that is sampleID
PCOA$X.SampleID <- IDs


#merge map and PCOA by SampleID - select both lines and run together - won't work otherwise

PCOA <- merge(PCOA, map, by = 'X.SampleID')


#### Plot PCOA ####

#instantiate shape and color palettes 

# shape_pal <- c(1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8)
# shape_pal2 <- c(1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3)
# shape_pal3 <- c(15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18)
colpal<- colorRampPalette(brewer.pal(12,"Set3"))
colpal<- sample(colpal(length(unique(PCOA$UserName))))

id <- gsub("_", " ",id)


# make a variable for self/other to use for plotting
PCOA <- PCOA %>% mutate(User = ifelse(UserName == id, "self", "other"))

# sort data frame so that the values matching id are at the bottom and plotted last
PCOA <- PCOA[order(PCOA$User),]
 
ggplot(PCOA, aes(x = PC1, y = PC2, group = User)) +
  geom_point(aes(fill = UserName, color = ifelse(PCOA$UserName %in% id == TRUE, "you","z")),
             alpha=1, pch = 21, size = 5) +
  scale_fill_manual(breaks = id, values = colpal) +
  scale_color_manual(values = c("black", "grey")) +
  scale_shape_manual(values = shape_pal3) +
  stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95) +
  labs(title=paste0("Beta diversity of your gut microbiome - PCoA -  ", id)) +
  guides(fill = guide_legend(override.aes = list(alpha=1)), size = FALSE, col = FALSE) +
  #guides(size = FALSE, fill = FALSE) +
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        legend.position = "bottom")



  
```
**Figure 6** is a PCoA plot. A PCoA plot summarizes variability within a given dataset by producing a set of uncorrelated axes. Essentially, the PCoA plot can be utilized to interpret similarity of data points -- data points closer to one another are more similar to one another, while points further away from each other are more dissimilar. Each data point in this plot represents a unique sample collected one of the participants in this study. Samples from the same study participant are the same color. 

The points corresponding to your microbiome are outlined in black and are highlighted by the smaller ellipse. When viewing this plot, keep in mind that it is a 2-dimensional representation of a 3-dimensional plot. So if your points look like they are all in the same place, they might actually extend into or out of the page!

\newpage
##Figure 7
```{r,fig.width=15, fig.height=15, echo=FALSE, warning = FALSE, message=FALSE}
grad1<- c("#fee6ce", "#fdae6b", "#e6550d")
grad2<- c("#deebf7", "#9ecae1", "#3182bd")
grad3<- c("#e5f5e0", "#a1d99b", "#31a354")
grad4<- c("#efedf5", "#bcbddc", "#756bb1")

Bact <- ggplot(PCOA, aes(x = PC1, y = PC2, group = User)) +
  geom_point(aes(fill = Bacteroidetes, color = ifelse(PCOA$UserName %in% id == TRUE, "you","z")),
             alpha=1, pch = 21, size = 5) +
  scale_fill_gradientn(colors = grad1) +
  scale_color_manual(values = c("black", "dark grey")) +
  #scale_shape_manual(values = shape_pal3) +
  stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95) +
  labs(title="Relative Abundance of Bacteroidetes") +
  guides(fill = guide_legend(override.aes = list(alpha=1)), size = FALSE, col = FALSE) +
  #guides(size = FALSE, fill = FALSE) +
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        legend.position = "bottom")

Firm <- ggplot(PCOA, aes(x = PC1, y = PC2, group = User)) +
  geom_point(aes(fill = Firmicutes, color = ifelse(PCOA$UserName %in% id == TRUE, "you","z")),
             alpha=1, pch = 21, size = 5) +
   scale_fill_gradientn(colors = grad2) +
  scale_color_manual(values = c("black", "dark grey")) +
  #scale_shape_manual(values = shape_pal3) +
  stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95) +
  labs(title="Relative Abundance of Firmicutes") +
  guides(fill = guide_legend(override.aes = list(alpha=1)), size = FALSE, col = FALSE) +
  #guides(size = FALSE, fill = FALSE) +
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        legend.position = "bottom")

Prot <- ggplot(PCOA, aes(x = PC1, y = PC2, group = User)) +
  geom_point(aes(fill = Proteobacteria, color = ifelse(PCOA$UserName %in% id == TRUE, "you","z")),
             alpha=1, pch = 21, size = 5) +
  scale_fill_gradientn(colors = grad3) +
  scale_color_manual(values = c("black", "dark grey")) +
  #scale_shape_manual(values = shape_pal3) +
  stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95) +
  labs(title="Relative Abundance of Proteobacteria") +
  guides(fill = guide_legend(override.aes = list(alpha=1)), size = FALSE, col = FALSE) +
  #guides(size = FALSE, fill = FALSE) +
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        legend.position = "bottom")


Actin <- ggplot(PCOA, aes(x = PC1, y = PC2, group = User)) +
  geom_point(aes(fill = Actinobacteria, color = ifelse(PCOA$UserName %in% id == TRUE, "you","z")),
             alpha=1, pch = 21, size = 5) +
  scale_fill_gradientn(colors = grad4) +
  scale_color_manual(values = c("black", "dark grey")) +
  #scale_shape_manual(values = shape_pal3) +
  stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95) +
  labs(title="Relative Abundance of Actinobacteria") +
  guides(fill = guide_legend(override.aes = list(alpha=1)), size = FALSE, col = FALSE) +
  #guides(size = FALSE, fill = FALSE) +
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        legend.position = "bottom")



plot_grid(Bact, Firm, Prot, Actin, nrow = 2, ncol = 2 )

```

**Figure 7** shows the relative abundance of key phyla within your gut microbiome compared with the rest of subjects in the study. As with Figure 6, your points are shown outlined in black and are surrounded by the smaller ellipse. 


\newpage

##Figure 8
```{r,fig.width=15, fig.height=15, echo=FALSE, warning = FALSE, message=FALSE}

grad1<- c("#fee6ce", "#fdae6b", "#e6550d")
grad2<- c("#deebf7", "#9ecae1", "#3182bd")
grad3<- c("#e5f5e0", "#a1d99b", "#31a354")
grad4<- c("#efedf5", "#bcbddc", "#756bb1")

Bactr <- ggplot(PCOA, aes(x = PC1, y = PC2, group = User)) +
  geom_point(aes(fill = Bacteroides, color = ifelse(PCOA$UserName %in% id == TRUE, "you","z")),
             alpha=1, pch = 21, size = 5) +
  #scale_fill_gradientn(colors = grad1) +
  scale_color_manual(values = c("black", "dark grey")) +
  #scale_shape_manual(values = shape_pal3) +
  stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95) +
  labs(title="Relative Abundance of Bacteroides ") +
  guides(fill = guide_legend(override.aes = list(alpha=1)), size = FALSE, col = FALSE) +
  #guides(size = FALSE, fill = FALSE) +
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        legend.position = "bottom")

Para <- ggplot(PCOA, aes(x = PC1, y = PC2, group = User)) +
  geom_point(aes(fill =Parabacteroides, color = ifelse(PCOA$UserName %in% id == TRUE, "you","z")),
             alpha=1, pch = 21, size = 5) +
  #scale_fill_gradientn(colors = grad2) +
  scale_color_manual(values = c("black", "dark grey")) +
  #scale_shape_manual(values = shape_pal3) +
  stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95) +
  labs(title="Relative Abundance of Parabacteroides") +
  guides(fill = guide_legend(override.aes = list(alpha=1)), size = FALSE, col = FALSE) +
  #guides(size = FALSE, fill = FALSE) +
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        legend.position = "bottom")

Alis <- ggplot(PCOA, aes(x = PC1, y = PC2, group = User)) +
  geom_point(aes(fill = Alistipes, color = ifelse(PCOA$UserName %in% id == TRUE, "you","z")),
             alpha=1, pch = 21, size = 5) +
  #scale_fill_gradientn(colors = grad3) +
  scale_color_manual(values = c("black", "dark grey")) +
  #scale_shape_manual(values = shape_pal3) +
  stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95) +
  labs(title="Relative Abundance of Alistipes ") +
  guides(fill = guide_legend(override.aes = list(alpha=1)), size = FALSE, col = FALSE) +
  #guides(size = FALSE, fill = FALSE) +
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        legend.position = "bottom")


Prevo <- ggplot(PCOA, aes(x = PC1, y = PC2, group = User)) +
  geom_point(aes(fill = Prevotella, color = ifelse(PCOA$UserName %in% id == TRUE, "you","z")),
             alpha=1, pch = 21, size = 5) +
  #scale_fill_gradientn(colors = grad4) +
  scale_color_manual(values = c("black", "dark grey")) +
  #scale_shape_manual(values = shape_pal3) +
  stat_ellipse(color = "dark grey", type = "norm", linetype = 2, level = 0.95) +
  labs(title="Relative Abundance of Prevotella") +
  guides(fill = guide_legend(override.aes = list(alpha=1)), size = FALSE, col = FALSE) +
  #guides(size = FALSE, fill = FALSE) +
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(plot.title = element_text(size = 14, hjust = 0.5),
        legend.position = "bottom")


plot_grid(Bactr, Para, Alis, Prevo, nrow = 2, ncol = 2 )

```



\newpage
#Your blood draw data

The results of your blood draw at the start of the dietary supplementation period are shown in the table below along with the reference ranges. If any of your values fall outside the reference ranges you should contact your health care provider to discuss your results and have the tests rerun.


| Test                  | Result                              | Normal Range |
|-----------------------|-------------------------------------:|:--------------:|
| Cholesterol (mg/dL)   | `r subblood$Cholesterol.Baseline`   |  <200        |
| Triglycerides (mg/dL) | `r subblood$Trigs.Baseline`         |  <150        | 
| HDL (mg/dL)           | `r subblood$HDL.Baseline`           |  >60         |  
| LDL (mg/dL)           | `r subblood$LDL.Baseline`           |  <100        |   
| Glucose (mg/dL)       | `r subblood$Glu.Baseline`           | 70 - 100     |
