---
output: pdf_document
fig_caption: true
params: 
    set_title: "My Title!"
---


---
title: `r "Microbiome Diet Study Participant Report"`
---


Thank you for participating in the Knights Lab citizen science project: The Microbiome Diet Study. Included in this report is some high-level information about your daily dietary intake and your daily microbiome variation. This report does not provide any medical advice and is not intended to be used to diagnose disease. If you have questions about your report, or you would like access to your raw data, please contact the study coordinator Abby Cole at cole0463@umn.edu




Here are your average daily macronutrient nutritional intakes contrasted to recommended macronutrient nutritional intake levels:

&nbsp;



 

| Type                | Your Macronutrient Intake Proportions                     | Recommended Intake Range    |
|---------------------|-----------------------------------------------------------|-----------------------------|
| CALORIES (kcal)     | `r round(mean(submap$KCAL),2)` kilocalories               |                             |
| PROTEIN (g)         | `r round((mean(submap$PROT)/macro_totals)*100,2)`%        |   10 - 35%                  | 
| TOTAL FAT (g)       | `r round((mean(submap$TFAT)/macro_totals)*100,2)`%        |   20 - 35%                  |  
| CARBS (g)           | `r round((mean(submap$CARB)/macro_totals)*100,2)`%        |   45 - 65%                  |                                      

&nbsp;

These are your average daily micronutrient nutritional intakes contrasted to recommended micronutrient nutritional intake levels:


| Type                | Your Average                   | Study Average                          |Recommended Daily Allowances(Male/Female)|
|---------------------|--------------------------------|----------------------------------------|-----------------------------------------|          
| FOLATE (ug)         | `r round(mean(submap$FOLA),2)` | `r round(mean(map$FOLA,na.rm=TRUE),2)` |                 400                     |
| CHOLINE (mg)        | `r round(mean(submap$CHOLN),2)`| `r round(mean(map$CHOLN,na.rm=TRUE),2)`|                550 / 425                |
| CALCIUM (mg)        | `r round(mean(submap$CALC),2)` | `r round(mean(map$CALC,na.rm=TRUE),2)` |                 1000 *                  |
| SODIUM (mg)         | `r round(mean(submap$SODI),2)` | `r round(mean(map$SODI,na.rm=TRUE),2)` |                 1500 *                  |
| POTASSIUM (mg)      | `r round(mean(submap$POTA),2)` | `r round(mean(map$POTA,na.rm=TRUE),2)` |                 4700                    |
| MAGNESIUM (mg)      | `r round(mean(submap$MAGN),2)` | `r round(mean(map$MAGN,na.rm=TRUE),2)` |               400 / 310 *               | 
| IRON (mg)           | `r round(mean(submap$IRON),2)` | `r round(mean(map$IRON,na.rm=TRUE),2)` |                8 / 18  *                |
| SELENIUM (ug)       | `r round(mean(submap$SELE),2)` | `r round(mean(map$SELE,na.rm=TRUE),2)` |                 55                      |
| ZINC (mg)           | `r round(mean(submap$ZINC),2)` | `r round(mean(map$ZINC,na.rm=TRUE),2)` |                11 / 8  *                |
| VITAMIN B12 (ug)    | `r round(mean(submap$VB12),2)` | `r round(mean(map$VB12,na.rm=TRUE),2)` |                 2.4                     |
| VITAMIN A (ug)      | `r round(mean(submap$VARA),2)` | `r round(mean(map$VARA,na.rm=TRUE),2)` |               900 / 700                 |
| VITAMIN C (mg)      | `r round(mean(submap$VC),2)`   | `r round(mean(map$VC,na.rm=TRUE),2)`   |                90 / 75                  |
| VITAMIN D (ug)      | `r round(mean(submap$VITD),2)` | `r round(mean(map$VITD,na.rm=TRUE),2)` |                 15 *                    |
| VITAMIN E (mg)      | `r round(mean(submap$ATOC),2)` | `r round(mean(map$ATOC,na.rm=TRUE),2)` |                 15                      |
| VITAMIN K (ug)      | `r round(mean(submap$VK),2)`   | `r round(mean(map$VK,na.rm=TRUE),2)`   |               120 / 90                  |


&nbsp;

The above tables show your average daily intakes of key macro and micro nutrients during the study period. For your reference also shown here are the overall average for the other participants in the study and the recommended intake levels by gender. Values marked with asterisks are indicative of recommended values that may vary for individuals over 30 may be higher or lower, visit https://ods.od.nih.gov/Health_Information/Dietary_Reference_Intakes.aspx for a comphrenhensive breakdown of dietary reference intakes by age and gender.

<!-- MicroNutrients -->

```{r, echo = F, warning=FALSE, message=FALSE,  fig.height=10, fig.width=10,fig.cap='Figure 1 visualizes day to day variation in your consumption of micronutrients viewed as a percentage of total kilocalorie intake. Protein intakes are abbreviated as "PROT", Carbohydrate intakes are abbreviated as "CARB", and Total Fat intakes are abbreviated as "TFAT" in the figure legend'}


long_plot <- submap %>% select(X.SampleID, KCAL, PROT, TFAT, CARB, UserName)
long_plot$PROT <- long_plot$PROT*4/long_plot$KCAL
long_plot$TFAT <- long_plot$TFAT*9/long_plot$KCAL
long_plot$CARB <- long_plot$CARB*4/long_plot$KCAL


long_plot <- melt(long_plot, id.vars = c("X.SampleID", "UserName", "KCAL"), 
                  measure.vars = c("PROT", "TFAT", "CARB"), 
                  variable.name = "Macro", value.name = "PCT")


#instantiate color brewer and axis tick mark amount
colpal <- colorRampPalette(brewer.pal(9,"Set3"))(37)
colpal <- sample(colpal,37,replace = FALSE)
tick_label <- seq(1:17)

# plot kcal filled with % protein, fat, and carb
ggplot(long_plot, aes(x = X.SampleID, y = KCAL*PCT, fill = Macro)) +
  geom_bar(stat = "identity") +
  facet_grid(.~UserName, scales = "free") +
  theme_classic() +
  scale_fill_manual(values=colpal)+
  #scale_fill_brewer(palette = colpal) +
   scale_x_discrete(labels=c(tick_label))+
  labs(fill = "% Macronutrient", y = "Daily reported intake (Kcal)", x = "Study Day Number") +
  theme(
        #axis.title.x =element_text(family = "Helvetica"), axis.title.y =element_text(family = "Helvetica"),
        plot.title = element_text(hjust = 0.5))+
ggtitle("Daily Micronutrient Consumption as a percentage of Kilocalorie Intake")

``` 

<!-- Daily Fiber Intake -->



```{r, warning=F, message=F, echo=FALSE, fig.cap=paste("Figure 2 visualizes your Fiber intake on a day-to-day basis. You had an average fiber consumption of",round(mean(submap$FIBE),2), "grams per day.The red line visualizes the recommended level of fiber consumption in grams, while the gold line visualizes your average daily fiber intake throughout the course of the study. Add blurb pertaining to microbiome importance")}

#instantiate color brewer and axis tick mark amounts
colpal<- colorRampPalette(brewer.pal(9,"Set3"))(37)
colpal<- sample(colpal,37, replace = FALSE)
tick_label <- seq(1:17) 

#Study average for individual's fiber intake
#fiber_average <-round(mean(submap$FIBE),2)

p <- ggplot(submap,aes(x = StudyDayNo,y = FIBE)) + 
     geom_bar(aes(fill = colpal[1]),stat = "identity",position = "dodge") + scale_x_discrete(labels = abbreviate)+
      geom_hline(size=.75,col="red",yintercept = 25) + #Recommended fiber level
      geom_hline(size=.75,col="gold",yintercept = fiber_average)  #Recommended fiber level
#fill = studydayno
p+guides(fill=F) +
  theme_classic(base_family = "Helvetica")+
   theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_discrete(labels=c(tick_label))+
   scale_fill_manual(values = colpal)+
  labs(x="Study Day Number", y= "Fiber levels in grams")+
  ggtitle("Daily Fiber Intake")
```

<!-- Microbiome Daily Relative Abundance -->

```{r, warning=F, message=F, echo=FALSE, fig.cap='Figure 3 depicts the most abundant bacterial species within your gut per each day of the study. The "<2.5% abundance" column represents a sum of bacterial species that individually account for less than 2.5% of the bacterial population within your gut microbiome.'}
# make ggplot bar chart of top 10 most abundant species per day
  
#melt subtaxasp to get our dataframe in the long format for future usage
  
  meltdf<- melt(subtaxasp)
  
  #merge to get access to Day var
  mergedf<- merge(x=meltdf, y=map, by.x = "variable", by.y= "X.SampleID", all.x=TRUE)
  
  #convert our dataframe species (rn) column to a character
  mergedf$rn <- as.character(mergedf$rn) 
  
  #series of gsub commands meant to neaten and clarify legend content
  mergedf$rn <- gsub(".*s__", "", mergedf$rn)
  mergedf$rn <- gsub("\\[", "",mergedf$rn)
  mergedf$rn <- gsub("\\]", "",mergedf$rn)
  mergedf$rn <- gsub("_", " ",mergedf$rn)
  
  
  #create small abundance category
  mergedf$rn[mergedf$value < 0.025] <- "<2.5% abundance"
  mergedf <- mergedf %>% group_by(StudyDayNo, rn) %>% summarise(newvalue = sum(value))
  
  
  #instantiate color brewer and axes tick mark amount
  colpal<- colorRampPalette(brewer.pal(7,"Set3"))(25)
  #colpal<- sample(colpal,37, replace = FALSE)
  tick_label <- seq(1:17)

  
ggplot(mergedf, aes(x = StudyDayNo, y = newvalue, fill = rn)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(drop = FALSE,labels=c(tick_label)) +
   scale_fill_manual(values =colpal)+
  theme_classic(base_family = "Helvetica") +
  theme(strip.text.y = element_text(angle = 0, size = 8, face = "italic", family = "Helvetica"),
        axis.text.x = element_text( hjust = 1, family = "Helvetica"),
        plot.title =  element_text(hjust = 0.5, family = "Helvetica"),
        strip.background = element_rect(color = "grey")) +
  guides(fill = guide_legend(reverse = FALSE, 
                             keywidth = 1, 
                             keyheight = 1, 
                             ncol = 1,
                             title= "Bacterial Species")) +
 labs(y="Relative Abundance",x="Study Day Number")+
  ggtitle("Main species within your gut per day")





```


<!-- Food abundance bar charts -->
```{r, warning=F, message=F, echo=FALSE, fig.cap='Figure 4 depicts the relative abundance of primary food groups you consumed on a day to day basis. Each color is representative of a unique food group as annotated on the figure legend. '}

# make ggplot bar chart of top 10 most abundant species per day
#melt sub_foodsp to get our dataframe in the long format for future usage (differs from melted_food)
  melt_food<- melt(sub_foodsp)
  
  #merge to get access to Day var (differs from merged_food)
  merge_food<- merge(x=melt_food, y=map, by.x = "variable", by.y= "X.SampleID", all.x=TRUE)
  
  #rename columns to get rid of the .x following UserName and StudyDayNo
  
  colnames(merge_food)[4] <- "UserName" #Change col name from UserName.x to UserName - compatibility purposes
  colnames(merge_food)[5] <- "StudyDayNo" #Change col name from StudyDayNo.x to StudyDayNo - compatibility purposes
  
  #convert our dataframe species (rn) column to a character
  merge_food$rn <- as.character(merge_food$rn) 
  
  #series of gsub commands meant to neaten and clarify legend content
  merge_food$rn <- gsub(".*s__", "", merge_food$rn)
  merge_food$rn <- gsub("\\[", "",merge_food$rn)
  merge_food$rn <- gsub("\\]", "",merge_food$rn)
  merge_food$rn <- gsub("_", " ",merge_food$rn)
  merge_food$rn <- gsub("L1", "",merge_food$rn)
  
  
  #instantiate color brewer and axes tick mark amount
  colpal<- colorRampPalette(brewer.pal(12,"Set3"))(500)
  colpal<- sample(colpal,500, replace = FALSE)
  tick_label <- seq(1:17)

  
ggplot(merge_food, aes(x = StudyDayNo, y = value, fill = rn)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(drop = FALSE,labels=c(tick_label)) +
   scale_fill_manual(values =colpal)+
  theme_classic(base_family = "Helvetica") +
  theme(strip.text.y = element_text(angle = 0, size = 8, face = "italic", family = "Helvetica"),
        axis.text.x = element_text( hjust = 1, family = "Helvetica"),
        plot.title =  element_text(hjust = 0.5, family = "Helvetica"),
        strip.background = element_rect(color = "grey")) +
  guides(fill = guide_legend(reverse = TRUE, 
                             keywidth = 1, 
                             keyheight = 1, 
                             ncol = 1,
                             title= "Food Group")) +
 labs(y="Relative Abundance",x="Study Day Number")+
  ggtitle("Main foods within your gut per day")






```




<!--Subject Day-to-Day Variation in Alpha Diversity of the Microbiome-->


```{r, warning=F, message=F, echo=FALSE, fig.cap='Figure 5 details how the bacterial diversity exhibited within your gut changes on a daily basis. This diversity catalogued within the gut microbiome is known as alpha diversity, and the metric utilized is the Shannon index of alpha diversity. The Shannon index accounts for both abundance and eveness of bacterial species present within the gut microbiome. '}
 #transpose
   subtaxaalpha <- t(subtaxaalpha)
 
  #calculate alpha diversity and assign to variable alphad
   alphad <- diversity(subtaxaalpha, index = "shannon", MARGIN = 1, base = exp(1))
   alphad <- as.data.frame(alphad)
   alphad <- rownames_to_column(alphad, var = "X.SampleID")
   
   #merge melted df with map to get access to StudyDay variable
   merged_alpha<- merge(x=alphad, y=map, all.x=TRUE)
    tick_label <- seq(1:17)
   colpal<- colorRampPalette(brewer.pal(9,"Set3"))(37)

#make gg plot line graph of subject alpha diversity per day
ggplot(merged_alpha, aes(x=StudyDayNo, y=alphad, group = UserName)) +
  geom_point() +
  theme(
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  theme_classic()+
  geom_smooth(aes(fill=colpal[1])) +
  scale_x_discrete(drop = FALSE,labels=c(tick_label)) +
  labs(x= "Study Day Number",y="Alpha Diversity (Shannon)")+
  guides(fill=FALSE)+
  ggtitle(" Day-to-Day Variation in Alpha Diversity of the Gut Microbiome")
  
```


<!--Subject Day-To-Day Variation in Beta Diversity of the Microbiome-->

```{r,fig.width=10, fig.height=10, echo=FALSE, fig.cap='Figure 6 is a PCoA plot. A PCoA plot summarizes variability within a given dataset by producing a set of uncorrelated axes. Essentially, the PCoA plot can be utilized to interpret similarity of data points -- data points closer to one another are more similar to one another, while points further away from eachother are more dissimilar. Each data point in this plot represents a subject gut microbiome at a particular time during the multiple days of sample collection and testing. The shape correspinding to your subject ID (given above) represents data points pertaining to your gut microbiome. Shapes corresponding to your subject ID have been enlarged for your viewing convenience. '}
#create PCOA plot of beta diversities of pertinent subjects


betad<-vegdist(betataxa, method="bray", binary=FALSE, diag=FALSE, upper=FALSE,
        na.rm = FALSE) 

#betad as matrix 
betad <- as.matrix(dist(betad))

#column name is "IDs"
IDs <-colnames(betad)

# Run the pcoa() function on the beta diversity table, and store the vectors generated as a dataframe 
PCOA <- data.frame(pcoa(betad)$vectors)

#create vector with placeholders

new_names <- rep("",ncol(PCOA))

#create for loop which gives column names (currently named "axis") the name of PC1, PC2, PC3.. so forth
#Each column for the PCOA matrix is a new PC (number).

for(i in 1:ncol(PCOA)){
  new_names[i] <- paste("PC",i,sep="")  
}

#code above uses a for loop - from the range of 1 to the number of columns in the PCOA
#matrix, the for loop accesses the new name variable and names it PC + i (which is 1,2,3  depending on column
#as indicated by for loop - for loop increases in increments of 1 (i goes up by 1) and goes thru columns)

#new_names (PC1, PC2, etc..) replace the current colnames
names(PCOA) <- new_names

#creates column that is sampleID
PCOA$X.SampleID <- IDs


#merge map and PCOA by SampleID - select both lines and run together - won't work otherwise

PCOA <- merge(PCOA, map, by = 'X.SampleID')


PCOA <- merge(PCOA, map, by = 'UserName')


#### Plot PCOA ####


#instantiate shape and color palettes 

shape_pal <- c(1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8)
shape_pal2 <- c(1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3,1,2,3)
shape_pal3 <- c(15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18,15,19,17,18)
colpal<- colorRampPalette(brewer.pal(9,"Set3"))(37)
id <- gsub("_", " ",id)


#use ggplot to plot
ggplot(PCOA) +
  geom_point(aes(x=PC1,y=PC2, col=UserName, shape=UserName,
  size=ifelse(PCOA$UserName %in% id == TRUE,15,5)),alpha=.060, stroke=3) + #gives axes and item of focus
  labs(title=paste0(" Day-To-Day Variation in Beta Diversity of the Gut Microbiome - PCOA -  ", id) ) +
   guides( colour = guide_legend(override.aes = list(size=5,alpha=1))) +
  scale_shape_manual(values=shape_pal3) +
  scale_color_manual(values=colpal)+
  theme_classic(base_family = "Helvetica") +                                                                      
  theme(legend.position = "bottom", plot.title = element_text(size = 18, hjust = 0.5))

 


  
```


